Lyman-alpha spectra from rotating outflows

### 2018-03-09 paper/paper.tex
edit

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
distributions of galaxies.
We perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these two
features impact the morphology of the Lyman-$\alpha$ emission line.
One of the most notorious results is that rotation increases the
detected flux at the line's center in comparison with the
non-rotational case. 
Rotation also induces a line broadening and a dependency with the
viewing angle.
We present a semi-analytic model that modifies the spectra of
non-rotational that manages to reproduce the main features of these
Monte Carlo simulations. 
\end{abstract}

@@

The interpretation of the Lyman-$\alpha$ emission in galaxies is
key to understand their evolutionary processes \citep{PartridgePeebles}.
Recent improvements in instrumentation have revolutionized the kind of
studies that can be performed on Lyman-$\alpha$ emitting galaxies
(LAEs.)
For instance, it is now possible to infer detailed kinematic maps for
nearby galaxies. 
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

These kinematic maps show that lyman alpha emitting galaxies have two 
overimposed velocity patterns: coherent rotation plus and a random
motions around the mean value provided by rotation. 
Depending on the system under study, one of the two features can be
more important, or they can be found in equal measure. 
These kinematic features play an important role in shaping the \lya.

@@

We know that rotation alone does have an impact on the \lya line morphology
and we get the same influence in it that previous works. Together with the 
outflows, we found that the resulting spectra is equivalent to a non-rotating
galaxy with outflows that undergoes a Doppler Shift in velocity, weighted by 
amount of photons that if rotating would project their escape in $\hat{k}$'s
direction. The formulas to calculate this shift are:

\begin{equation}
V' = V + \vec{v}_{\rm rot}\cdot\hat{k} = V + V_{\rm rot} (x k_{x} - y k_{y})
\label{eq:shift_x}
\end{equation}

where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by 
outflows and $\vec{v}_{\rm rot} = V_{\rm rot} (x\hat{i} - y\hat{j})$.

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
distributions of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these two
features impact the morphology of the Lyman-$\alpha$ emission line.
One of the most notorious results is that rotation increases the
detected flux at the line's center in comparison with the
non-rotational case. 
Rotation also induces a line broadening and a dependency with the
viewing angle.
We present a semi-analytic model that modifies the spectra of
non-rotational that manages to reproduce the main features of these
Monte Carlo simulations. 
As an application example we contraint the rotation parameters to the
observed Lyman-$\alpha$ profile of a high redshift LAE, providing the
first a constraint on the rotational velocity of a high redshift
galaxy based on its integrated \lya emission.
\end{abstract}

@@

The interpretation of the Lyman-$\alpha$ emission in galaxies is
key to understand their evolutionary processes, specially in
star-forming, low-dust galaxies \citep{PartridgePeebles}.
Recent improvements in instrumentation have revolutionized the kind of
studies that can be performed on Lyman-$\alpha$ emitting galaxies
(LAEs.)
For instance, it is now possible to infer detailed kinematic maps for
nearby galaxies. 
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

These kinematic maps show that LAEs have two  overimposed velocity
patterns: coherent rotation plus and a random motions around the mean
value provided by rotation.  
Depending on the system under study, one of the two features can be
more important, or they can be found in equal measure.  
These kinematic features play an important role in shaping the \lya.

@@

Rotation alone does have an impact on the \lya line morphology
and we get the same influence in it that previous works. 
Together with the  outflows, we found that the resulting spectra is
equivalent to a non-rotating galaxy with outflows that undergoes a
Doppler Shift in velocity, weighted by  amount of photons that if
rotating would project their escape in $\hat{k}$'s direction:

\begin{equation}
V' = V + \vec{v}_{\rm rot}\cdot\hat{k} 
\label{eq:shift_x}
\end{equation}
%
where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by 
outflows.
